Supervised Learning for Psychological Attention to High School Students

### 2020-11-18 article/article.tex
[ article ] finished results for NN

\author{\IEEEauthorblockN{Juan S. Cárdenas-Rodríguez}
  \IEEEauthorblockA{\textit{Department of Mathematical Sciences} \\
    \textit{EAFIT University}\\
    Medellín, Colombia \\
    jscardenar@eafit.edu.co} \and \IEEEauthorblockN{David Plazas}
  \IEEEauthorblockA{\textit{Department of Mathematical Sciences} \\
    \textit{EAFIT University}\\
    Medellín, Colombia \\
    dplazas@eafit.edu.co} }

@@

\section{Methodology}\label{sec:meth}
All the code developed for this work was implemented in Julia 1.5.3. and
can be found in the authors'
\href{https://github.com/juanscr/ai-works}{GitHub repository}.

@@

Additionally, it is well-known that the bound for the training dataset size is
given by
\begin{equation}
  N\geq \dfrac{\ln(2)}{2\epsilon^2}\left[(2^k-1)(1+\log_2 m) + 1 + \ln\left(
  \dfrac{1}{\delta}\right)\right],
\end{equation}
but again, the size of the dataset is fixed. Thus, $\epsilon$ was calculated by
fixing the tree depth $k=4$ and setting $\delta=0.7$ yielding a value of
$\epsilon=0.5381$, which is a suitable value for the classifier error.

@@

@@

\subsection{Neural Networks}
\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-gradients.pdf}
        \caption{Gradients for each layer.}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-error.pdf}
        \caption{Average error energy per epoch.}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-roc.pdf}
        \caption{ROC curve.}
    \end{subfigure}
    \caption{Learning curves for $\mathcal{L}=[1,2,3]$ network $\eta=0.9$
    (worst).}
    \label{fig:NN-worst}
\end{figure*}

@@

\section{Results for Embedded Data}
This section presents the same results above presented but for the data
embedded with the TSNE algorithm prior to training.

>>>>

\author{\IEEEauthorblockN{Juan Sebastián Cárdenas-Rodríguez}
  \IEEEauthorblockA{\textit{Department of Mathematical Sciences} \\
    \textit{EAFIT University}\\
    Medellín, Colombia \\
    jscardenar@eafit.edu.co} \and \IEEEauthorblockN{David Plazas}
  \IEEEauthorblockA{\textit{Department of Mathematical Sciences} \\
    \textit{EAFIT University}\\
    Medellín, Colombia \\
    dplazas@eafit.edu.co} }

@@

\section{Methodology}\label{sec:meth}
All the code developed for this work was implemented in Julia 1.5.3. and
can be found in the authors'
\href{https://github.com/juanscr/ai-works}{GitHub repository}, although the
plots were generated using Python 3.8.6.

@@

Additionally, it is well-known that the bound for the training dataset size is
given by
\begin{equation}
  N\geq \dfrac{\ln(2)}{2\epsilon^2}\left[(2^k-1)(1+\log_2 m) + 1 + \ln\left(
  \dfrac{1}{\delta}\right)\right],
\end{equation}
but again, the size of the dataset is fixed. Thus, $\epsilon$ was calculated by
fixing the tree depth $k=4$ and setting $\delta=0.3$ yielding a value of
$\epsilon=0.5381$, which is a suitable value for the classifier error.

@@

The results in the following sections were obtained by using the 6-dimensional
dataset. All the procedures were also applied to the embedded 2-dimensional
data (via TSNE) but the results are presented in Appendix \ref{app:emb}.

@@

\subsection{Neural Networks}
Recall that only the worst and 2 best NN are kept for presenting the results.
The worst NN (in the sense of the objective function) was trained with
$\eta=0.9$ and it has a $\mathcal{L}=[1,2,3]$ architecture in the hidden layers.
The learning curves for this MLP are presented in Fig. \ref{fig:NN-worst}.
This architecture is expected to perform poorly, at least for the given number
of epochs, since it is attempting to summarize all 6 inputs in a one dimensional
value for the first hidden layer and then casting this value into higher
dimensions. It is unkown for the authors if this machine can learn the data at
all, since the behavior of the backpropagation algorithm is unpredictable as
epochs advance. Note from Fig. \ref{subfig:grad-worst-NN}, the gradients are
stable in the epoch horizon defined, from Fig. \ref{subfig:error-worst-NN} that
the average energy error seems to reach a stable point, but different from 0,
finally, from Fig. \ref{subfig:roc-worst-NN}, the ROC curve and AUC confirms the
poor performance of this classifier. Finally, the Sensitivity-Specificity
results (see Table \ref{tab:L123}) show that this NN has adequate values for the
specificity but deplorable values for the sensitivity.

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-gradients.pdf}
        \caption{Gradients for each layer.}
        \label{subfig:grad-worst-NN}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-error.pdf}
        \caption{Average error energy per epoch.}
        \label{subfig:error-worst-NN}
    \end{subfigure}
    \begin{subfigure}[b]{0.32\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/1-2-3-0.9-roc.pdf}
        \caption{ROC curve.}
        \label{subfig:roc-worst-NN}
    \end{subfigure}
    \caption{Learning curves for $\mathcal{L}=[1,2,3]$ network $\eta=0.9$
    (worst).}
    \label{fig:NN-worst}
\end{figure*}

Furthermore, the second-best NN was obtained with $\mathcal{L}=[2]$ and the best
with $\mathcal{L}=[3]$, both trained with $\eta=0.9$. It is believed that this
MLP obtained better results since the number of weights to optimize was
relatively low and, hence, the backpropagation algorithm converges quickly
(taking into consideration) the elevated learning rate. The training curves for
these two models are presented respectively in Figs. \ref{fig:NN-2best} and
\ref{fig:NN-best}. Note that both networks show rich dynamics on the gradients
through the training stage, the average error energy seems to be descending into
0 and both ROC curves are close to the perfect constant value of 1. The
Sensitivity-Specificity values also show good results on all three sets for the
$\mathcal{L}=3$ architecture.

@@

\section{Results for Embedded Data}\label{app:emb}
This section presents the same results above presented but for the data
embedded with the TSNE algorithm prior to training.
